Reproducible Research Assignment 1
========================================================

## Loading and preprocessing the data
Data is loaded into R and date field is recast from factor to date.  
```{r}
## Data loaded into R
data <- read.csv("activity.csv",header=TRUE)

## Date Recast to date format using lubridate package
library(lubridate)
data$date <- ymd(data$date)
```

Initial exploration identifies several days with all NA values.  It is assumed the measuring device was not worn on these days.  These will be dealt with later in the analysis.
```{r fig.width=7, fig.height=6}
## explore null values by day
data$isNull <- is.na(data$steps)
library(ggplot2)
qplot(date,isNull,data=aggregate(isNull ~ date, data=data,sum))
```

Initial analysis also identifies several days where nearly all values are 0.  It is assumed the measuring device was not worn on these days either, but these will be left in the data set.

```{r fig.width=7, fig.height=6}
## explore zero values by day
data$isZero <- data$steps==0
library(ggplot2)
qplot(date,isZero,data=aggregate(isZero ~ date, data=data,sum))
```

The data now appears ready for analysis.
```{r}
## Pull initial stats
str(data)
head(data)
```

## What is mean total number of steps taken per day?
Steps are aggregated by day into a new data frame.  A plot of steps per day is generated, and the mean and median of steps per day is calculated.
```{r fig.width=7, fig.height=6}
## Prep Data frame
stepsPerDay <- aggregate(steps ~ date,data=data,sum)

## Examine Data
head(stepsPerDay)
str(stepsPerDay)

## Plot Historgram
qplot(date,steps,data=stepsPerDay)

## Calculate Mean
mean(stepsPerDay$steps)

## Calculate Median
median(stepsPerDay$steps)
```

## What is the average daily activity pattern?
Steps are averaged across 5 minute intervals into a new data frame.  A plot is generated of average steps by interval, and the maximum steps per interval is calculated.
```{r fig.width=7, fig.height=6}
## Prep Data Frame
avgStepsInt <- aggregate(steps ~ interval,data=data,mean)

## Examine Data
head(avgStepsInt)
str(avgStepsInt)

## Generate Plot
qplot(interval,steps,data=avgStepsInt,geom=c("smooth","point"))

## Pull maximum average steps and interval for maximum steps
max(avgStepsInt$steps)
avgStepsInt[avgStepsInt$steps==max(avgStepsInt$steps),"interval"]
```

## Imputing missing values
As noted above, there appears tobe missing data in the dataset.
```{r}
## Calculates # of missing observations
sum(is.na(data$steps))
```

To impute missing values, we will use the mean value for that interval from the rest of the data set.
```{r}
## Create new data frame with imputed data set
library(plyr)
names(avgStepsInt)=c("interval","avgSteps")
aggregateData <- join(data,avgStepsInt,by="interval")
for (i in 1:nrow(aggregateData)) {
  if (is.na(aggregateData[i,"steps"])==TRUE){
    aggregateData[i,"steps"] <- aggregateData[i,"avgSteps"]
  }
}
newData <- aggregateData[,c("steps","interval","date")]

# Examine new data
str(newData)
head(newData)
```

The analysis from the second part is recreated on the new data set.
```{r fig.width=7, fig.height=6}
## Prep Data frame
stepsPerDay <- aggregate(steps ~ date,data=newData,sum)

## Examine Data
head(stepsPerDay)
str(stepsPerDay)

## Plot Historgram
qplot(date,steps,data=stepsPerDay)

## Calculate Mean
mean(stepsPerDay$steps)

## Calculate Median
median(stepsPerDay$steps)
```
The imputatation did not appear to have a meaningful impact the mean and median of the data set.

## Are there differences in activity patterns between weekdays and weekends?
We first create a new factor variable identifying a day as a weekday or weekend.
```{r}
for (i in 1:nrow(newData)) {
  if (weekdays(newData[i,"date"])=="Saturday" | weekdays(newData[i,"date"])=="Saturday"){
    newData[i,"dayOrEnd"] <- "Weekend"
  }
  else {
    newData[i,"dayOrEnd"] <- "Weekday"
  }
}
newData$dayOrEnd <- as.factor(newData$dayOrEnd)

## Examine data frame
summary(newData)
```

We now calculate the average steps per interval again, this time separating by weekday or weekend.  The results are examined and plotted.
```{r fig.width=7, fig.height=6}
## Prep Data Frame
avgStepsInt <- aggregate(steps ~ interval + dayOrEnd,data=newData,mean)

## Examine Data
head(avgStepsInt)
str(avgStepsInt)

## Generate Plot
qplot(interval,steps,data=avgStepsInt,geom="line",facets=.~dayOrEnd)